---
title: "Module_3_app1"
author: "Priya Shaji"
date: "3/2/2020"
output: html_document
runtime: shiny
---

> Question 1:
As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.


```{r}
## Import necessary package

library(dplyr)
library(shiny)
library(ggplot2)
library(rsconnect)

## Reading data

## Skipping 879 rows, as it shows error as: duplicate 'row.names' are not allowed`

data <- read.csv("https://raw.githubusercontent.com/Shetura36/Data-608/master/module%203/cleaned-cdc-mortality-1999-2010-2-ver2.csv",skip=879, nrows=49806,header=TRUE, sep=",", strip.white = TRUE, quote = "\"'")

## Remove rows that do not have values for State attribute
data <-data[!(data$State==''),]

## Selecting complete cases only 

data <- data[complete.cases(data), ]

## Set column names

colnames(data) <- c('Disease', 'State', 'Year', 'Deaths', 'Population', 'Crude.Rate')

## Change data type for `Disease` and `Crude.Rate` attributes

data$Disease <- as.character(data$Disease)
data$Crude.Rate <- as.numeric(data$Crude.Rate)

## Extract only disease categories from string

data$Disease <- substr(data$Disease, regexpr('>', data$Disease)+1, nchar(data$Disease))

## Subset for year 2010 only

data_2010 <- data[data$Year == 2010, ]

## Calculate national average of crude mortality rate

avg_data <- data_2010 %>% group_by(Disease)  %>% summarise(total_population = sum(Population)) %>% inner_join(data_2010, by="Disease")
avg_data$Weighted_CrudeRate <- (avg_data$Population/avg_data$total_population) * avg_data$Crude.Rate

## Add disease_avg to data_2010

data_2010 <- 
  avg_data %>% group_by(Disease) %>% summarise(disease_avg = sum(Weighted_CrudeRate)) %>% inner_join(data_2010, by="Disease")

## National average for each disease

avg_2010 <- as.data.frame(unique(data_2010 %>% group_by(Disease) %>% select(Disease, disease_avg)))

```



```{r}

## UI
ui <- fluidPage(
  sidebarLayout(
    
    # Input
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "disease", 
                  label = "Select disease:",
                  choices = unique(data_2010$Disease),
                  selected = "Neoplasms",
                  width = '500px')
      
    ),
    
    # Output:
    mainPanel(
      plotOutput(outputId = "bargraph", width = "100%")
    )
  )
)

## Define server function required to create the scatterplot-
server <- function(input, output, session) {
  
  #This is an event reactive element that responds to a specific event (in this case a button clicked)
  #and sets the valueExpr to a specific value (in this case the plot title)
  ordered <- eventReactive(
    eventExpr = input$disease, 
    valueExpr = {data_2010 %>% filter(Disease==input$disease) %>% arrange(desc(Crude.Rate))},
    ignoreNULL = FALSE
  )
  
  #Create bar graph
  output$bargraph <- renderPlot({ggplot(data = ordered(), aes(x=reorder(State,Crude.Rate), y=Crude.Rate)) + 
      geom_bar(stat="identity", width=0.7, color="#1F3552", fill="steelblue", 
               position=position_dodge()) +
      geom_text(aes(label=round(Crude.Rate, digits=2)), hjust=1.3, size=3.0, color="white") + 
      coord_flip() + 
      ggtitle("Crude Mortality for Selected Disease for 2010") +
      xlab("") + ylab("") + 
      theme_minimal()}, height = 1000, width = 600)
}

## Create a Shiny app object
shinyApp(ui = ui, server = server)

```

