
---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

1) Import the required packages

```{r}
library(ggplot2)
library(kableExtra)
library(dplyr)

```

If we analyze the growth rates, min is 0.340 and it maximum growth rate is 421.480. Let's see how many companies experianced growth rates of 50 or higher:

```{r}

inc %>% dplyr::filter(Growth_Rate >= 50) %>% summarise(n = n())

```


Therefore, there were 59 companies with growth greater than or equal to 50.

And below is the list of those companies:

```{r}
kable(inc %>% dplyr::filter(Growth_Rate >= 50)) %>% kable_styling()
```


2) Now let's see how many distinct companies exists in our dataset:

```{r}
kable(inc %>% dplyr::group_by(Industry) %>% dplyr::summarise(n =n()) %>% arrange(desc(n))) %>% kable_styling()
```

3) Let us calculate the median revenue:

```{r}
inc %>% dplyr::summarise(min=min(Revenue), median=median(Revenue), max=max(Revenue))
```


4) Let's calculate number of distinct cities:

```{r}
cities <- inc %>% group_by(City) %>% summarise((n=n()))
nrow(cities)
```

Below are the top 10 companies and the cities they are located in:

```{r}
kable(inc %>% group_by(City) %>% summarise(n=n()) %>% arrange(desc(n)) %>% top_n(10)) %>% kable_styling()
```




5) Now let's see the employee minimum and maximum range in the companies listed in the dataset.

```{r}
kable(inc %>% dplyr:: summarise(min=min(Employees, na.rm = TRUE), median=median(Employees, na.rm = TRUE), max=max(Employees, na.rm =TRUE))) %>% kable_styling()
```


6) Number of states that are distinct in the given dataset:

```{r}
distinct_states <- inc %>% group_by(State) %>% summarise(n=n())
nrow(distinct_states)
```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
# Answer Question 1 here

## distribution of companies in the dataset by State
companies_byState <- inc %>% group_by(State) %>% summarise(n=n()) %>% arrange(desc(n))

plt1 <- 
  ggplot(data = companies_byState[1:52,], aes(x=reorder(State,n), y=n)) + 
  geom_bar(stat="identity", width=0.5, color="#AA4371", fill="steelblue", 
           position=position_dodge()) +
    #geom_text(aes(label=round(n, digits=2)), hjust=1.3, size=3.0, color="white") + 
    coord_flip() + 
    scale_y_continuous(breaks=seq(0,700,100)) + 
    ggtitle("Disbribution by State") +
    xlab("") + ylab("") + 
    theme_minimal()

## Plotting the graph

plt1

```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}

# Answer Question 2 here

## state with the 3rd most companies in the data set

kable(inc %>% group_by(State) %>% summarise(n=n()) %>% arrange(desc(n)) %>% top_n(3)) %>% kable_styling()

```


As shown ablove, the state with third most companies in the dataset is New York.

Let's evaluate the company cases:

```{r}
company_cases <- inc[complete.cases(inc),]

```


Now, let's find the median number of employees in each industry for NY state.

```{r}
kable(company_cases%>%filter(State=='NY')%>%group_by(Industry)%>%summarise(min=min(Employees),median=median(Employees),max=max(Employees),var=var(Employees)) %>% arrange(desc(var))) %>% kable_styling()
```


The data above shows the min, median, and max number of employees for each industry in NY. It is ordered from highest to lowest variability.
 
 
In order to show the median number of employees, a box plot could be plotted.The plot will also display range of data and outliers. Number of distinct industries are 25. Let's use the table above to show that companies that have higher variability in employee number is ones with higher maximum number of employees.

Below, the industries are grouped together.

```{r}
Business_Products_Services <- c('Business Products & Services')
Consumer_Products_Services <- c('Consumer Products & Services')
group_2 <- c('Travel & Hospitality', 'Human Resources', 'IT Services', 'Software')
group_3 <- c('Security', 'Media', 'Financial Services',  'Environmental Services', 'Food & Beverage')
group_4 <- c('Energy', 'Telecommunications', 'Manufacturing', 'Health', 'Construction')
group_5 <- c('Advertising & Marketing', 'Education', 'Engineering', 'Logistics & Transportation', 'Retail')
group_6 <- c('Insurance', 'Real Estate', 'Computer Hardware', 'Government Services')
```


Creating box plots for the respective groups:

```{r}
plt_Business_Products_Services <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% Business_Products_Services), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE)

plt_Consumer_Products_Services <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% Consumer_Products_Services), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE)

plt_group_2 <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% group_2), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE)

plt_group_3 <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% group_3), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE)

plt_group_4 <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% group_4), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE)

plt_group_5 <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% group_5), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE) 

plt_group_6 <- ggplot(company_cases %>% filter(State=='NY' & Industry %in% group_6), aes(x = Industry, y = Employees)) + 
        coord_flip() + 
        geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=1, notch=FALSE)
```


```{r}
plt_Business_Products_Services
```


```{r}
plt_Consumer_Products_Services
```



Box plots for remaining industries: x-axis scale for each group is different.

The box plots for ‘Business Products & Services’ and ‘Consumer Products & Services’ came out as very small. The outlier data is causing the box plot of these 2 industries to flatten out.


```{r}
plt_group_2
plt_group_3
plt_group_4
plt_group_5
plt_group_6
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here

revenue_perEmployee <- 
company_cases %>% group_by(Industry) %>% summarise(count=n(), total_revenue=sum(Revenue), total_employees=sum(Employees), revenue_perEmployee=total_revenue/total_employees) %>% arrange(desc(revenue_perEmployee))

kable(revenue_perEmployee) %>% kable_styling()

```


The code below plots the revenue per employee as a bar chart sorted by revenue per employee from highest to lowest. 

A second bar chart plot is generated that shows the distribution of companies by industry sorted by revenue per employee from highest to lowest (same order as the first plot).


```{r}
plt_revenue_perEmployee_a<- ggplot(data=revenue_perEmployee, aes(x=reorder(Industry,-revenue_perEmployee), y=revenue_perEmployee)) +
     geom_bar(stat="identity", fill="steelblue") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
     ggtitle("Revenue Per Employee by Industry") + 
     ylab("Revenue Per Employee") + 
     xlab("")

plt_revenue_perEmployee_b <- ggplot(data=revenue_perEmployee, aes(x=reorder(Industry,-revenue_perEmployee), y=count)) +
     geom_bar(stat="identity", fill="steelblue") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
     ggtitle("Distribution of Companies by Industry") + 
     ylab("Revenue Per Employee") + 
     xlab("")

plt_revenue_perEmployee_a
plt_revenue_perEmployee_b
```

